Determinism, Complexity, and Predictability in Computer Performance

### 2014-07-14 wpe.tex
fix the lower limit of a sum
second round of revisions.  awaiting today's conversation before can finish.

From a different point of view, we can consider the information (as
measured by the Shannon entropy) contained in a single observable, say
the \emph{present}, of the system. This information can be partitioned
into two components: the information shared with past
observations---e.g., the mutual information between the past and
present---and the information in the present that is not contained in
the past (aka ``the conditional entropy of the present given the
past'').  The first part is known as the \emph{redundancy}, as it is
information in the present that is also in the past.  The second part
is the aforementioned \emph{Shannon entropy rate}.  We conjecture that 
the more redundancy in a signal, the more predictable it is.  
Redundancy is critical to prediction, in general.  And the specific 
\emph{form} of the redundancy dictates whether a particular prediction 
method will work well or poorly on the corresponding signal.  A linear 
method cannot capture or make use of nonlinear redundancy, for 
instance.  This issue is central to the claims in this paper and the 
discussion in the following section.

@@

Here we will be utilizing a variation of the basic permutation entropy
technique, the \emph{weighted permutation entropy} (WPE), which was
introduced in~\cite{fadlallah2013}.  The intent behind the weighting
is to correct for observational noise that is larger than the trends
in the data, but smaller than the larger scale features.  Consider,
for example, a signal that switches between two fixed points and
contains some additive noise. The PE is dominated by the noise about
the fixed points driving it to $\approx 1$, which in some sense hides
the fact that the signal is actually quite structured.  In this
situation, the terms of the weighted permutation entropy are dominated
by the switching rather than the stochastic fluctuations. To
accomplish this, the \emph{weight} of a permutation is taken into
account:
\begin{align*}
  w(x_{i+1}^\ell) = \frac{1}{\ell} \sum_{j = i}^{i+\ell}
                      \left( x_j - \bar{x}_{i+1}^\ell \right)^2
\end{align*}
where $x_{i+1}^\ell$ is a sequence of values $x_{i+1}, \ldots,
x_{i+\ell}$, and $\bar{x}_{i+1}^\ell$ is the arithmetic mean of
those values.

>>>>

From a different point of view, we can consider the information (as
measured by the Shannon entropy) contained in a single observable, say
the \emph{present}, of the system. This information can be partitioned
into two components: the information shared with past
observations---e.g., the mutual information between the past and
present---and the information in the present that is not contained in
the past (aka ``the conditional entropy of the present given the
past'').  The first part is known as the \emph{redundancy}, as it is
information in the present that is also in the past.  The second part
is the aforementioned \emph{Shannon entropy rate}.  It seems obvious
that the more redundancy in a signal, the more predictable it is.
Redundancy is critical to prediction, in general.  And the specific
\emph{form} of the redundancy dictates whether a particular prediction
method will work well or poorly on the corresponding signal.  A linear
method cannot capture or make use of nonlinear redundancy, for
instance.  This issue---which has been explored extensively using
clean data and/or systems where the generating process is known, but
not so much with empirical data from unknown systems---is central to
the claims in this paper and the discussion in the following section.

@@

Here we will be utilizing a variation of the basic permutation entropy
technique, the \emph{weighted permutation entropy} (WPE), which was
introduced in~\cite{fadlallah2013}.  The intent behind the weighting
is to correct for observational noise that is larger than the trends
in the data, but smaller than the larger-scale features.  Consider,
for example, a signal that switches between two fixed points and
contains some additive noise. The PE of such a signal will be
dominated by the noise about the fixed points, driving it to $\approx
1$---which in some sense hides the fact that the signal is actually
quite structured.  In this situation, the terms of the weighted
permutation entropy are dominated by the switching rather than the
stochastic fluctuations. To accomplish this, the \emph{weight} of a
permutation is taken into account:
\begin{align*}
  w(x_{i+1}^\ell) = \frac{1}{\ell} \sum_{j = i + 1}^{i+\ell}
                      \left( x_j - \bar{x}_{i+1}^\ell \right)^2
\end{align*}
where $x_{i+1}^\ell$ is a sequence of values $x_{i+1}, \ldots,
x_{i+\ell}$, and $\bar{x}_{i+1}^\ell$ is the arithmetic mean of
those values.
